Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation

### 2020-11-11 xrb2/paper.tex
added my orcid
Updates to flame propagation section of XRB2 paper

@@

\begin{deluxetable}{lrrr}
	\tablecaption{\label{table:500Hz_runs} Temperature properties of the 500~Hz simulations.}
	\tablehead{\colhead{run} & \colhead{$\delta T$ (K)} & \colhead{$T_\mathrm{hi}$ (K)} & \colhead{$T_\mathrm{star}$ (K)}} 
	\startdata
	{\tt flame\_wave\_500}      & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
	{\tt flame\_wave\_500\_hot} & $1.2\times 10^9$ & $2\times 10^8$ & $2\times 10^8$ \\
	{\tt flame\_wave\_500\_3E8} & $1.1\times 10^9$ & $3\times 10^8$ & $3\times 10^8$ \\
	{\tt flame\_wave\_500\_4E8} & $10^9$ & $4\times 10^8$ & $4\times 10^8$ \\
	\enddata
	%
\end{deluxetable}

@@

different rotation rates

@@

Figure~\ref{fig:flame_speeds} shows the position of the flame as a
function of time.  Defining the position is difficult, so we \MarginPar{need explanation here}

\begin{figure}[t]
	\centering
	\plottwo{speeds_1_100th_max.pdf}{speeds_1_1000th_max.pdf}
	\caption{\label{fig:flame_speeds} Plots of flame front position vs.\ time for the 500 Hz and 1000 Hz runs. The left-hand plot shows the results when tracking the $\frac{1}{100} \max[\enucdot]$ contour, while the right-hand plot shows the results for the $\frac{1}{1000} \max[\enucdot]$ contour.}
\end{figure}

>>>>

% Approximately proportional to, from StackOverflow
\def\app#1#2{%
	\mathrel{%
		\setbox0=\hbox{$#1\sim$}%
		\setbox2=\hbox{%
			\rlap{\hbox{$#1\propto$}}%
			\lower1.1\ht0\box0%
		}%
		\raise0.25\ht2\box2%
	}%
}
\def\appropto{\mathpalette\app\relax}

@@

\begin{deluxetable}{lrrr}
	\tablecaption{\label{table:500Hz_runs} Temperature properties of the 500~Hz simulations.}
	\tablehead{\colhead{run} & \colhead{$\delta T$ (K)} & \colhead{$T_\mathrm{hi}$ (K)} & \colhead{$T_\mathrm{star}$ (K)}} 
	\startdata
	{\tt flame\_wave\_500}      & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
	% Do we want the 100 Hz run in here as well, for completeness (and so we have a name to reference)?
	{\tt flame\_wave\_500\_2E8} & $1.2\times 10^9$ & $2\times 10^8$ & $2\times 10^8$ \\
	{\tt flame\_wave\_500\_3E8} & $1.1\times 10^9$ & $3\times 10^8$ & $3\times 10^8$ \\
	{\tt flame\_wave\_500\_4E8} & $10^9$ & $4\times 10^8$ & $4\times 10^8$ \\
	\enddata
	%
\end{deluxetable}

@@

different rotation rates. 250 Hz failed to ignite, could indicate that another source of confinement (e.g.\ magnetic fields) would need to take
over at lower rotation rates to allow a burst to occur, at least for our current initial model. We've seen that increasing rotation rate can
increase nuclear reaction rate already

@@

\subsection{Flame Propagation}

For the purpose of measuring the flame propagation speed and acceleration, we track the position of each of our flames
as a function of time. We define the position in terms of a particular $\enucdot$ value, as we did in \citet{flame_wave1}.
To recapitulate: we first reduce the 2D $\enucdot$ data for each simulation run to a set of 1D radial profiles by averaging
over the vertical coordinate. After averaging, we take our reference $\enucdot$ value to be some fraction of the global $\enucdot$
maximum across all of these profiles. Since the flames in our simulations propagate rightward, we then search the region of
each profile to the right of the local $\enucdot$ maximum for the point where the $\enucdot$ first drops below this reference
value. This point gives us the location of our flame front.

In \citet{flame_wave1}, we used $0.1\%$ of the global $\enucdot$ maximum for our reference value. For the high temperature
unboosted flames, however, we found that the $\enucdot$ profiles failed to reach that small a value across the domain at most times,
which prevented us from getting reliable position measurements. We therefore use $1\%$ of the global $\enucdot$ maximum in this paper
rather than $0.1\%$. This is still sufficiently small that our measurements are not overly sensitive to turbulence and other local fluid
motions (the issue with simply tracking the local maximum), but allows us to avoid the pitfall encountered by the $0.1\%$ metric.

\begin{figure}[t]
	\centering
	\plotone{flame_speeds_rotation.pdf} % Not certain what style we want for the plot here
	\caption{\label{fig:flame_speeds_1} Plots of flame front position vs.\ time for the 500 Hz and 1000 Hz runs. The dashed lines
		show quadratic least-squares fits to the data for $t \gtrsim 20~\mathrm{ms}$.}
\end{figure}

\begin{figure}[t]
	\centering
	{\color{red} Placeholder for the flame speed plot that shows temperature dependence}
	\caption{\label{fig:flame_speeds_2} Plots of flame front position vs.\ time for the three 500 Hz runs with
		$T_{star} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
		for $t \gtrsim 20~\mathrm{ms}$.}
\end{figure}

Figures~\ref{fig:flame_speeds_1} and \ref{fig:flame_speeds_2} give position vs.\ time for our runs, with the former showing dependence
on rotation rate and the latter on the neutron star temperature. As some of the runs (in particular the $\tt{flame\_wave\_3E8}$ run) exhibit
significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit
parameters. These results are given in Table~\ref{table:flame_speeds}. We also include the flame speed at $t = 100~\mathrm{ms}$ extrapolated
from the fit.

We found in \citet{flame_wave1} that the flame speed $s$ obeys:
\begin{equation}
	s \appropto L_R (\epsdot)^{\half} \propto \frac{(\epsdot)^{\half}}{\Omega}
\end{equation}
where $\epsdot$ is the specific energy generation rate and $L_R$ is the Rossby length. This was consistent with the results of
\citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length. As seen in
Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling with rotation rate from this set of runs. We observed earlier,
however, that nuclear reactions progress more quickly at the higher rotation rate. This results in a higher $\enucdot$ --- up to 4
to 6 times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the
increased Coriolis confinement. We also observe that 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
small deceleration at early times. This disparity could be due to the 500 Hz flame taking more time to develop
fully and reach higher burning rates, or a result of the difference in Coriolis force. \MarginPar{Do we have other ideas why?}

% It might be worth plotting peak enucdot as a function of time for each run

{\color{red} Maybe Nicole can talk a bit about the effect of the neutron star temperature here? I have a placeholder for the plot earlier, but
feel free to move it. What's in this section now is just an initial draft/outline.}

\begin{deluxetable}{lrrr}
	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each simulation. $s_0$ is the speed at $t = 0~\mathrm{ms}$, and $s_{100}$ is the speed at $t = 100~\mathrm{ms}$.}
	\tablehead{\colhead{run} & \colhead{$s_0$ (km s$^{-1}$)} & \colhead{$s_{100}$ (km s$^{-1}$)} & \colhead{$a$ (km s$^{-2}$)}} 
	\startdata
	{\tt flame\_wave\_500}      & $3.08 \pm 0.01$ & $3.62 \pm 0.02$ & $5.41 \pm 0.19$ \\
	{\tt flame\_wave\_1000}     & $3.41 \pm 0.01$ & $3.31 \pm 0.02$ & $-1.03 \pm 0.13$ \\
	{\tt flame\_wave\_500\_2E8} & $3.76 \pm 0.02$ & $\sim 3.93$ & $1.74 \pm 0.17$ \\
	{\tt flame\_wave\_500\_3E8} & $5.29 \pm 0.10$ & $\sim 41.0$ & $357.4 \pm 1.5$ \\
	\enddata
	%
\end{deluxetable}
